---
title: "Web APIS- New York Times"
author: "Jack Wright"
date: "10/23/2020"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE}
library(tidyverse)
library(jsonlite)
library(httr)
```

## Process

I am using the NY Times API for article searches. I want to search for articles about the New York Jets from the sports desk over the last year. I will be generating the API request programattically. 

## Creating the request for the GET() function in HTTR:

### Load api key

```{r}
location<-"c:\\password-files-for-r\\nytimes_keys.csv"
nytimes_keys<-read.csv(location)

base_url<-"https://api.nytimes.com/svc/search/v2/articlesearch.json?"
#main query
q<-"q=New+York+Jets"
#&news_desk=sports&begin_date=20200101&end_date=20201023
#elements of fq
key<-paste0("api-key=",nytimes_keys$api_key)
tag<-paste(q,key,sep="&")

url<-paste0(base_url,tag)

```


## Requesting the Data

I will use the GET() function from the httr package and check the status to see if it was successful

```{r}
jets_pull<-GET(url)

http_status(jets_pull)
```


### Inspecting the data

The request to the api is structured as a nested named list. I need to find out where the content I am interested in is located.

```{r}
#look at the names
names(jets_pull)

#i want the content, but its contents is just raw bytes
glimpse(jets_pull$content)
#data is in bytes, so convert to text
jets_content<-fromJSON(rawToChar(jets_pull$content))

#after some checks I found where the data I am interested is located
names(jets_content$response$docs)

```



## Convert to Data frame

Since the data is structured as a list, I will convert it to a data frame.

```{r}
df_jets<-data.frame(jets_content$response$docs)

```


#Tidy the Data

The headline column for this data frame is a nested data frame. I will need to unnest it in order to select the main headline.

Then I will create a new dataframe suitable of looking at what my API request returned
```{r}
#unnest headline and put it in its own data frame
df_headline<-unnest(df_jets$headline)

output<-data.frame("main_headline"=df_headline$main,"abstract"=df_jets$abstract,"date"=df_jets$pub_date)

output%>%
  mutate(ymd=as.Date(date))%>%
  select(-date)
```


## Conclusions

My pull was only for the first page of results of jets articles (the most recent). I could create a function that allows me to add a pagination facet, allowing me to cycle through the results pages and pull more data. I could have also added more facets to my data frame, like only pulling from the sports desk.