---
title: "Homework 5"
author: "Jack Wright"
date: "9/24/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Load Packages
```{r message=FALSE}
library(tidyverse)
library(RCurl)

```

##Get data from github using RCurl package

```{r}
url<-"https://raw.githubusercontent.com/JackJosephWright/data/master/hw5.csv?token=AQZFVAMURDMZ6UXUQUADSEK7NVAUI"



dat<-read.csv(url)


```


#Tidying data
 lets take a look
 
 First, I want to remove the divider row from the table and fill out the carrier column
 
```{r}
#rename a couple columns for ease of use
names(dat)[1]<-"carrier"
names(dat)[2]<-"status"




#remove blank row
dat1<-dat%>%
  rowwise()%>%
  filter(!is.na(Los.Angeles))

#fill carrier column, fill() is not working I will dig deeper if i have time

dat1$carrier[2]<-"ALASKA"
dat1$carrier[4]<-"AM WEST"



```
 
 
 In order for data to be tidy, every column must be a variable and every row must be an observation. On first glance, the destinations look like values of a "destination" column and their values look like "count".

the "X column looks like values of a "carrier" variable
```{r}
#converting "destinations"  to destination column and their values to "count"
carrier_status<-dat1%>%
  pivot_longer(c("Los.Angeles":"Seattle"),names_to="destination",values_to="count")%>%
  arrange(destination)
  
```


## Data Analysis

I want to see which carrier supplied more on time flights. 

I will turn the counts into percentages, only display the on time results, because the delayed is redundant, and drop the "status and "count" columns

```{r}
carrier_stats<-carrier_status%>%
  group_by(carrier,destination)%>%
  mutate(on_time_rate=count/sum(count))%>%
  filter(status=="on time")%>%
  select(-status, -count)
head(carrier_stats)
```


I want to make a histogram

comparing each carrier at each city in a histogram

```{r}
carrier_stats%>%
  ggplot(aes(fill=carrier,y=on_time_rate,x=destination))+
  geom_bar(position="dodge",stat="identity")+coord_flip()

```


